---
title: "Introduction to R"
subtitle: "Data Visualization in R"
output: html_document
---

## How to follow along and participate

```{r chunk1}
library(package = "tidyverse")
library(package = "haven")
library(package = "gghalves")
```

## Importing data into R

```{r chunk2}

globalData2019 <- read_spss(file = "Pew Research Center Global Attitudes Spring 2019 Dataset WEB.sav")
```

## Data cleaning

```{r chunk3}

globalData2019clean <- globalData2019 %>%
  filter(country %in% c(13, 17, 30, 32)) %>% 
  select(country, SEX, AGE, BETTER_GENDER, COUNTRY_SATIS, 
         ECON_SIT, LADDER_NOW) %>% 
  zap_labels() %>%
  mutate(AGE = na_if(AGE, 98)) %>% 
  mutate(AGE = na_if(AGE, 99)) %>% 
  mutate(LADDER_NOW = na_if(LADDER_NOW, 98)) %>% 
  mutate(LADDER_NOW = na_if(LADDER_NOW, 99)) %>% 
  mutate(country = recode_factor(country,
                             '13' = 'Israel',
                             '17' = 'Lebanon',
                             '30' = 'Tunisia',
                             '32' = 'Ukraine')) %>%
  mutate(SEX = recode_factor(SEX,
                             '1' = 'Male',
                             '2' = 'Female')) %>% 
  mutate(BETTER_GENDER = recode_factor(BETTER_GENDER,
                                    '1' = 'Men',
                                    '2' = 'Women',
                                    '3' = 'Same/both equally',
                                    '8' = NA_character_,
                                    '9' = NA_character_)) %>% 
  mutate(COUNTRY_SATIS = recode_factor(COUNTRY_SATIS,
                                       '1' = 'Satisfied',
                                       '2' = 'Dissatisfied',
                                       '8' = NA_character_,
                                       '9' = NA_character_)) %>% 
  mutate(ECON_SIT = recode_factor(ECON_SIT,
                                  '1' = 'Very good',
                                  '2' = 'Somewhat good',
                                  '3' = 'Somewhat bad',
                                  '4' = 'Very bad',
                                  '8' = NA_character_,
                                  '9' = NA_character_)) 
```


## Check your work

```{r chunk4}

summary(object = globalData2019clean)
```

## Basic bar chart

```{r chunk10}

globalData2019clean %>%
  ggplot(aes(x = SEX)) +
  geom_bar() 
```

## Fancy bar chart

```{r chunk11}

globalData2019clean %>%
  ggplot(aes(x = SEX)) +
  geom_bar(fill = "deeppink") +
  labs(x = "Sex",
       y = "Frequency",
       title = "Sex of 4058 participants in a 2019 survey of people from Israel,\nLebanon, Tunisia, and Ukraine.") +
  theme_minimal()
```

## Fancy bar chart with colors

```{r chunk12}

globalData2019clean %>%
  ggplot(aes(x = SEX, fill = SEX)) +
  geom_bar() +
  labs(x = "Sex",
       y = "Frequency",
       title = "Sex of 4058 participants in a 2019 survey of\npeople from Israel,\nLebanon, Tunisia, and Ukraine.") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 14)
```

## Super fancy bar chart 

```{r chunk13}

globalData2019clean %>%
  ggplot(aes(x = SEX, fill = SEX)) +
  geom_bar() +
  labs(x = "Sex",
       y = "Frequency",
       title = "Sex of 4058 participants in a 2019 survey of\npeople from Israel,\nLebanon, Tunisia, and Ukraine.") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 14, base_family = "serif") +
  theme(legend.position = "none") 
```

## You try it!

Make a bar chart of the `BETTER_GENDER` responses from survey participants.

Edit the code below to: 

* Use `BETTER_GENDER` on the x-axis and as the fill variable in the `ggplot()` layer

* Change the color of the bars by replacing "Set2" with the name of another palette (hint: palette names can be found by using the Help tab and entering scale_fill_brewer)

* Add labels for the axes and a title inside the quote marks (see codebook for more about the question)

* Replace `theme_minimal()` with another theme, you can find them by typing `theme_` into the Help tab

* Play with the `base_size = ` by trying different numbers

When you have completed the changes, click the little green arrow on the top right of the code chunk to run the code.


```{r chunk14, eval = FALSE}

globalData2019clean %>%
  ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(x = "",
       y = "",
       title = "") +
  scale_fill_brewer(palette = "") +
  theme_minimal(base_size = , base_family = "serif") +
  theme(legend.position = "none") 
```


## Stacked bar plot

```{r chunk15}

globalData2019clean %>% 
  drop_na(BETTER_GENDER) %>% 
  ggplot(aes(x = BETTER_GENDER, fill = SEX)) +
  geom_bar() +
  labs(x = "Who has better life",
       y = "Frequency",
       title = "Who has a better life by sex according to 4058\nparticipants in a 2019 survey of people from Israel,\nLebanon, Tunisia, and Ukraine.") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal(base_size = 12, base_family = "mono") 
```

## Grouped bar plot

```{r chunk16}

globalData2019clean %>% 
  drop_na(BETTER_GENDER) %>% 
  ggplot(aes(x = BETTER_GENDER, fill = SEX)) +
  geom_bar(position = "dodge") +
  labs(x = "Who has better life",
       y = "Frequency",
       title = "Who has a better life by sex according to 4058 participants\nin a 2019 survey of people from Israel,\nLebanon, Tunisia, and Ukraine.",
       fill = "Survey participant sex") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal(base_size = 12, base_family = "serif") 
```


## You try it!

* Create a grouped bar plot with `COUNTRY_SATIS` as the x-axis and `SEX` as the fill color 

* Drop the `NA` values for `COUNTRY_SATIS`

* Add good labels for the x-axis, y-axis, title (may need the codebook for this)

* Format it using a palette you like  

* Choose a different theme and try it (hint: type theme_ into the help to get the names of other themes to try)

* Fix the legend title so it is "Participant sex"

```{r chunk16b, eval=FALSE}

globalData2019clean %>% 
  drop_na() %>% 
  ggplot(aes(x = , fill = )) +
  geom_bar(position = "dodge") +
  labs(x = "",
       y = "",
       title = "",
       fill = "") +
  scale_fill_brewer(palette = "") +
  theme_()
```


## Answer to You try it!

```{r chunk17}

globalData2019clean %>% 
  drop_na(COUNTRY_SATIS) %>% 
  ggplot(aes(x = COUNTRY_SATIS, fill = SEX)) +
  geom_bar(position = "dodge") +
  labs(x = "Satisfied or dissatisfied with the way things\nare going in our country ",
       y = "Frequency",
       title = "Country satisfaction by sex according to 4058 participants\nin a 2019 survey of people from Israel,\nLebanon, Tunisia, and Ukraine.", 
       fill = "Participant sex") +
  scale_fill_brewer(palette = "Paired") +
  theme_light() 

```

## Group percentages

```{r chunk19}

globalData2019clean %>% 
  group_by(COUNTRY_SATIS, SEX) %>% 
  count() %>% 
  group_by(COUNTRY_SATIS) %>% 
  mutate(perc = 100 * (n / sum(n))) 
```

## Grouped bar plot 

```{r chunk20}

globalData2019clean %>% 
  group_by(COUNTRY_SATIS, SEX) %>% 
  count() %>% 
  group_by(COUNTRY_SATIS) %>% 
  mutate(perc = 100 * (n / sum(n))) %>% 
  drop_na(COUNTRY_SATIS) %>% 
  ggplot(aes(x = COUNTRY_SATIS, y = perc, fill = SEX)) +
  geom_col(position = "dodge") +
  labs(x = "Satisfied or dissatisfied with the way things\nare going in our country ",
       y = "Percentage within satisfaction group",
       title = "Country satisfaction by sex according to 4058 participants\nin a 2019 survey of people from Israel,\nLebanon, Tunisia, and Ukraine.", 
       fill = "Participant sex") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() 
```

## Histogram

```{r chunk21}

# Histogram for age
globalData2019clean %>% 
  ggplot(aes(x = AGE)) +
  geom_histogram() 
```

## Fancy histogram

```{r chunk22}

# Histogram for age
globalData2019clean %>% 
  ggplot(aes(x = AGE, fill = SEX)) +
  geom_histogram(color = "white") +
  labs(x = "Participant age (years)",
       y = "Frequency",
       title = "Age distribution by sex for 4058 participants in a 2019 survey\nof people from Israel,\nLebanon, Tunisia, and Ukraine.") +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(rows = vars(SEX)) +
  theme_minimal() +
  theme(legend.position = "none") 
```

## Box plot

```{r chunk23}

globalData2019clean %>% 
  ggplot(aes(x = SEX, y = AGE)) +
  geom_boxplot() +
  labs(x = "Participant sex",
       y = "Age",
       title = "Age by sex for 4058 participants in a 2019 survey of people\nfrom Israel,\nLebanon, Tunisia, and Ukraine.") +
  theme_minimal() 
```

## Fancier box plot

```{r chunk24}

globalData2019clean %>% 
  ggplot(aes(x = SEX, y = AGE, color = SEX)) +
  geom_half_boxplot() +
  geom_half_point(alpha = .4) +
  labs(x = "Participant sex",
       y = "Age",
       title = "Age by sex for 4058 participants in a 2019 survey of people\nfrom Israel,\nLebanon, Tunisia, and Ukraine.") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Grouped box plot

```{r chunk25}

globalData2019clean %>% 
  ggplot(aes(y = LADDER_NOW, x = country, color = SEX)) +
  geom_boxplot() +
  labs(x = "Country",
       y = "Life ladder (0 = worst, 10 = best)",
       title = "Life ladder by sex and country for 4058 participants in a 2019 survey\nof people from Israel,\nLebanon, Tunisia, and Ukraine.",
       color = "Sex") +
  theme_minimal() 
```

## Rain cloud plot

```{r chunk26}

globalData2019clean %>% 
  ggplot(aes(x = SEX, y = AGE, color = SEX, fill = SEX)) +
  geom_half_violin(side = "r") +
  geom_half_point(alpha = .3, side = "l") +
  labs(x = "Participant sex",
       y = "Age",
       title = "Age by sex for 4058 participants in a 2019 survey of people from\nIsrael,\nLebanon, Tunisia, and Ukraine.") +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal() +
  coord_flip() +
  theme(legend.position = "none")
  
```

## Grouped rain cloud plot

```{r chunk27}

globalData2019clean %>% 
  ggplot(aes(y = LADDER_NOW, x = country, fill = SEX, color = SEX)) +
  geom_half_violin(side = "r") +
  geom_half_point(alpha = .3, side = "l") +
  labs(x = "Country",
       y = "Life ladder (0 = worst, 10 = best)",
       title = "Life ladder by sex and country for 4058 participants in a 2019 survey\nof people from Israel,\nLebanon, Tunisia, and Ukraine.",
       color = "Sex",
       fill = "Sex") +
  theme_minimal() +
  coord_flip() +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") 
  
```

## Grouped rain cloud plot with facets

```{r chunk28}

globalData2019clean %>% 
  ggplot(aes(y = LADDER_NOW, fill = SEX, color = SEX)) +
  geom_half_violin(side = "r") +
  geom_half_point(alpha = .3, side = "l") +
  labs(x = "Country",
       y = "Life ladder (0 = worst, 10 = best)",
       title = "Life ladder by sex and country for 4058 participants in a 2019 survey\nof people from Israel,\nLebanon, Tunisia, and Ukraine.",
       color = "Sex",
       fill = "Sex") +
  theme_minimal() +
  coord_flip() +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  facet_wrap(facets = vars(country))
  
```

## The final challenges...choose one or try them all!

* Make a bar graph showing percentage of participants in each category of ECON_SIT 

* Create a figure that shows grouped bar plots for GLOBAL_COMMUNITY grouped by country

* Create a figure that shows grouped bar plots for country and COUNTRY_SATIS

* Create a figure that shows grouped bar plots for COUNTRY_SATIS grouped by country and SEX

* Create a figure that shows the distribution of AGE by country and SEX

```{r chunk29}

```